# Correlation Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Exercise E1

### Part A

### Part B

### Part C

### Part D

## Exercise E2

## Exercise E3

### Part A

### Part B

## Exercise E4

### Part A

### Part B

## Exercise E5

A psychologist investigates different work groups in a company to see if there is a connection between the attractiveness of a work group (cohesion and well-being of its members) and the variance of its work performance.

Group attractiveness was measured using an ordinal scale from 1 = very high to 6 = very low. The variance in performance refers to the number of products completed by a work group on different days. The study produced the following result: 

```{r, echo=FALSE}
group <- LETTERS[1:7]
attractiveness <- c(1:2, 2:6)
variance <- c(32, 20, 40, 32, 64, 68, 60)
tbl <- tibble(group = group, attractiveness = attractiveness, variance = variance)
names(tbl) <- c("Work Group", "Attractiveness", "Variance in Performance")

kable(tbl)
```

### What is the statistical feature in this study?

### Calculate a suitable correlation coefficient.

### Interpret the calculated sign of this coefficient.

### Does the calculated value of this coefficient provide information about a cause-and-effect relationship between the two characteristics examined?
