---
title: "Statistics 1 for Economics"
author: "David Barkemeyer"
---

# Welcome {-}

This book contains the solutions to the exercises of the course Statistics 1 for
Economics at Technische Universitaet Berlin in the winter term 2020/21.

<!--chapter:end:index.Rmd-->

# (APPENDIX) Appendix {-}

# Frequencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Exercise A1

### Part A

The duration of study of 200 industrial engineers is examined. The following results were obtained:

```{r}
n <- 200
duration <- c(10, 11, 12, 13, 14, 15)
rel_frequency <- c(0.1, 0.1, 0.4, 0.2, 0.15, 0.05)
tbl <- tibble(duration = duration, rel_frequency = rel_frequency) %>%
  mutate(cum_rel_frequency = cumsum(rel_frequency))

tbl
```

#### What is the statistical feature? How is it scaled?

The feature is the duration study of an industrial engineer. It is interval-scaled.

#### Determine the absolute frequencies.

```{r}
tbl <- mutate(tbl, abs_frequency = n * rel_frequency)
tbl$abs_frequency
```

#### Determine the cumulative absolute frequencies.

```{r}
tbl <- mutate(tbl, cum_abs_frequency = cumsum(abs_frequency))
tbl$cum_abs_frequency
```

#### Plot frequencies and cumulative  (relative and absolute).

```{r}
# Relative frequencies
ggplot(tbl) +
  geom_col(aes(x = duration, y = rel_frequency))

# Relative cumulative frequencies
ggplot(tbl) +
  geom_point(aes(x = duration, y = cum_rel_frequency))

# Absolute frequencies
ggplot(tbl) +
  geom_col(aes(x = duration, y = abs_frequency))

# Absolute cumulative frequencies
ggplot(tbl) +
  geom_point(aes(x = duration, y = cum_abs_frequency))
```

#### What is the maximum duration of study of the 10%-fastest students?

10 semesters.

#### What is the minimum duration of study of the 80%-slowest students?

12 semesters.

#### What is the duration of study for exact 20% of all students?

13 semesters.

### Part B

Now we consider the feature $Y$: _duration of study_ with levels:

* *small* (less than 12 semesters)
* *medium* (12 semesters)
* *large* (more than 12 semesters)

#### How is the feature Y scaled?

It is ordinal-scaled.

#### Plot the absolute frequency function.

```{r}
# Function that converts duration to level
duration_to_level <- function(x) {
  if (x < 12) "small" else if (x == 12) "medium" else "large"
}
# Vectorize this function in order to work with vector input
duration_to_level <- Vectorize(duration_to_level)

# Prepare dataset
tbl <- select(tbl, duration, rel_frequency, abs_frequency) %>%
  mutate(level = duration_to_level(duration)) %>%
  group_by(level) %>%
  summarise(rel_frequency = sum(rel_frequency), abs_frequency = sum(abs_frequency))
tbl

# Plot
ggplot(tbl) +
  geom_col(aes(x = level, y = abs_frequency))
```

#### Does it make sense to use a cumulative distribution function for a nominal-scaled feature?

No.

## Exercise A2

The 100m times of 20 athletes were measured. They obtained the following cumulative distribution function:

```{r, echo=FALSE}
x <- c(10, 10.2, 10.6, 10.8, 11)
y <- c(0, 0.1, 0.6, 0.8, 1)
tbl <- tibble(x = x, y = y)
ggplot(tbl, aes(x = x, y = y)) +
  geom_line() +
  geom_point() +
  labs(x = "x", y = "F(x)")
```

### Draw the corresponding histogram

```{r}
break_left <- c(10, 10.2, 10.6, 10.8)
break_right <- c(10.2, 10.6, 10.8, 11)
f <- c(0.1 / 0.2, 0.5 / 0.4, 0.2 / 0.2, 0.2 / 0.2)
tbl <- tibble(break_left = break_left, break_right = break_right, f = f)

ggplot(tbl, aes(xmin = break_left, xmax = break_right, ymax = f)) +
  geom_rect(ymin = 0) +
  labs(x = "x", y = "f(x)")
```

### Which was the maximum time of the 80%-fastest athletes?

10.8 seconds.

### What is the assumption about the distribution of measured values within a group when determining the cumulative distribution function?

Uniform distribution within a group.

### Use the cumulative distribution function to determine which was the maximum time needed by the five fastest athletes.

```{r}
# Interpolation
t_max <- 10.2 + (0.25 - 0.1) / (0.6 - 0.1) * (10.6 - 10.2)
t_max
```

## Exercise A3

In a year tennis ace Boris B. participated in 40 tournaments all around the world. Every tournament consisted of six rounds:
* 1st Round
* 2nd Round
* 3rd Round (Round of 16)
* 4th Round (Round of 8)
* 5th Round (Semi-Final)
* 6th Round (Final)

Boris is not satisfied with his overall performance. Even though he reached the final two times and was six times only defeated in the semi-final, he lost ten times in the 1st Round and 16 times in the 2nd Round. He never dropped out in the 3rd Round.

### What is the statistical feature? How is it scaled?

The feature is the number of the highest round Boris reached in a tournament. It is interval-scaled.

### Determine absolute and relative frequencies as well as absolute and relative cumulative frequencies.

```{r}
x <- 1:6
abs_frequency <- c(10, 16, 0, 6, 6, 2)
tbl <- tibble(x = x, abs_frequency = abs_frequency)

tbl <- tbl %>%
  mutate(rel_frequency = abs_frequency / sum(abs_frequency)) %>%
  mutate(cum_abs_frequency = cumsum(abs_frequency)) %>%
  mutate(cum_rel_frequency = cumsum(rel_frequency))
tbl
```

### Draw the empirical relative distribution function.

```{r}
ggplot(tbl) +
  geom_col(aes(x = x, y = rel_frequency))
```

### At what percentage of tournaments Boris dropped out before the 3rd Round?

26 / 40.

### At what percentage of tournaments Boris reached at least the semi-finals?

8 / 40.

### At how many tournaments Boris played at least in the 2nd Round?

30.

### At most which round Boris dropped out in 80% of all tournaments?

4th Round.

### Which round Boris dropped out in exact 40% of all tournaments?

2nd Round.

### Give an interpretation of F(x) at x = 7.

$F(7) = 1$, because $F(x) = 1 \, \forall \, x >= 6$. That means it is impossible to ever reach the 7th Round as it does not exist.

## Exercise A4

100 students were asked how much time they spend for their studies per day. None of them spent more than 12 hours. 22 students spent more than six hours. 30% of all interviewed students were spending less than three hours, whereas 65% of students invest between three and eight hours per day.

### What is the statistical feature? How is it scaled?

The feature is the time a student spends for her studies per day in hours. It is ratio-scaled.

### Determine absolute, relative and cumulative frequencies. Group the data in four groups.

```{r}
n <- 100
break_left <- c(0, 3, 6, 8)
break_right <- c(3, 6, 8, 12)
# 22 students spent more than six hours -> 5 students spent more than 12 hours,
# 17 student spent between six and height hours
f <- c(0.3, 0.48, 0.17, 0.05)

tbl <- tibble(break_left = break_left, break_right = break_right, rel_frequency = f) %>%
  mutate(abs_frequency = rel_frequency * n) %>%
  mutate(cum_rel_frequency = cumsum(rel_frequency)) %>%
  mutate(cum_abs_frequency = cumsum(abs_frequency))

tbl
```

### Visualize the absolute frequencies.

```{r}
ggplot(tbl) +
  geom_rect(
    aes(
      xmin = break_left, 
      xmax = break_right, 
      ymin = 0, 
      ymax = abs_frequency / (break_right - break_left)
    ),
    fill = "gray",
    color = "black"
  ) +
  labs(x = "x", y = "h(x)")
```

### Why did you choose this kind of visualization?

I choose a histogram as it is a meaningful visualization for ratio-scaled data. Don't forget to divide by the group width. The area of a group is proportionate to its absolute frequency and not its height.

### Visualize the relative cumulative frequencies.

```{r}
x <- c(0, tbl$break_right)
cum_rel_frequency <- c(0, tbl$cum_rel_frequency)
tbl <- tibble(x = x, cum_rel_frequency = cum_rel_frequency)

ggplot(tbl, aes(x = x, y = cum_rel_frequency)) +
  geom_line() +
  geom_point() +
  labs(x = "x", y = "F(x)")
```

### Which assumption was made for your visualization?

The distribution within a group is uniform.

### Use your visualization to find out how many hours at most spent 50% of all students?

```{r}
# Interpolation
t_max <- 3 + (0.5 - 0.3) / (0.78 - 0.3) * (6 - 3)
t_max
```

### How many students work more than five hours per day for their studies?

```{r}
# F(5)
F_5 <- 0.3 + (5 - 3) / (6 - 3) * (0.78 - 0.3)
F_5

# Solution
(1 - F_5) * n
```

## Exercise A5

This exercise does not cover anything new and is therefore left as an exercise to the reader.

Ice cream seller Benedict wants to sell fresh ice cream to his customers every day. So that he always has the right quantities in stock, he counts the consumption of ice cream scoops on one day for 200 customers. Here it turns out:

* No customer is ordering two scoops or more than six scoops.
* 20% each wanted one scoop or at least five scoops.
* Less than three scoops were ordered by 45% of all customers.
* Ten times more customers wanted four or six scoops than five scoops.

### What is the statistical feature? How is it scaled?

### Determine absolute and cumulative absolute frequencies.

### Visualize the absolute frequencies.

### Visualize the relative cumulative frequencies.

### What percentage of customers ordered five or less scoops?

### How many scoops were bought from the 80% customers with the most scoops?

### How many scoops were bought from exact 35% of customers?
 

<!--chapter:end:01_A_Frequencies.Rmd-->

# Empirical Parameters

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Model Exercise B0 - Geometric mean {-}

Statistix would like to invest its savings of 1,000 EUR in a project. This project has a duration of five years and has the following annual interest rates:
```{r, echo=FALSE}
t <- c("t_0 = 0", "t_1 = 1", "t_2 = 2", "t_3 = 3", "t_4 = 4", "t_5 = 5")
i <- c("-", "10%", "12%", "35%", "27%", "4%")
x <- c(1000, 1100, 1232, 1663.2, 2112.26, 2196.75)
tbl <- tibble(t = t, i = i, x = x)

names(tbl) <- c("Time t [years]", "Interest rate i", "Year-end value [EUR]")

kable(tbl)
```

### Show that the interest rate i_2 for time t_2 is exactly 12%.

### Calculate the average interest rate per year based on the arithmetic mean.

### Check your result by calculating the average interest rate over 5 years on the starting capital.

### Now calculate the correct average annual interest rate.

### Check your result.

## Exercise B1

Mister Maier owns a garden gnome wholesale with three stores in Berlin, New York and Flensburg. At the end of the business year, he wants an overview of the business situation and therefore requests information in all three stores about the orders received within the last year.

### Part A

His Berlin store is providing the following data:

```{r, echo=FALSE}
x <- c("0 - 20", "20 - 50", "50 - 150", "150 - 300")
order <- c(15, 30, 45, 10)
tbl <- tibble(x = x, order = order)
names(tbl) <- c("Order value in 1000 EUR from ... to less than ...", "Number of orders")
kable(tbl)
```

```{r}
# Prerequisites
break_left <- c(0, 20, 50, 150)
break_right <- c(20, 50, 150, 300)
n_orders <- c(15, 30, 45, 10)
n_berlin <- sum(n_orders)

tbl <- tibble(break_left = break_left, break_right = break_right, n_orders = n_orders)

kable(tbl)
```

#### Determine relative cumulative frequencies.

```{r}
tbl <- tbl %>%
  mutate(rel_frequency = n_orders / 100) %>%
  mutate(cum_rel_frequency = cumsum(rel_frequency)) %>%
  print()
```

#### Determine lower and upper quartile and median. Draw a box plot.

```{r}
# Interpolation
x_25 <- 20 + (0.25 - 0.15) / (0.45 - 0.15) * (50 - 20)
x_50 <- 50 + (0.5 - 0.45) / (0.9 - 0.45) * (150 - 50)
x_75 <- 50 + (0.75 - 0.45) / (0.9 - 0.45) * (150 - 50)
x_25
x_50
x_75
```

#### Calculate the mean order value.

```{r}
tbl <- tbl %>%
  mutate(group_mean = (break_left + break_right) / 2) %>%
  print() %>%
  mutate(total_order_value = group_mean * n_orders) %>%
  print() %>%
  summarise(mean_order_value = sum(total_order_value) / sum(n_orders)) %>%
  print()

x_bar_berlin <- tbl$mean_order_value
```

#### What is your assumptions for this calculation?

Uniform distribution within a group.

#### What is the biggest possible difference between the mean of the original data and the mean of the grouped data

Half the group width, that means 75.000 EUR.

### Part B

The story in New York answers the request as follows:

$$95 \, \mathrm{orders} \, | \, \bar{x} = 60.000\$ \, | \, s^2 = 16.000.000\$^2$$

To improve comparability, Mister Maier wants to get his information in EUR.

#### Prerequisites {-}

```{r}
n_new_york <- 95
```

#### Calculate the corresponding values for an exchange rate of 1$ = 1.5 EUR.

```{r}
x_bar_dollar <- 60000
x_bar_new_york <- 60000 * 1.5

var_dollar <- 16e6
# 1st variant
sd_dollar <- sqrt(var_dollar)
sd_euro <- sd_dollar * 1.5
var_new_york <- sd_euro^2
# 2nd variant
var_new_york_2 <- var_dollar * 1.5^2

x_bar_new_york
var_new_york
var_new_york_2
```

### Part C

From Flensburg Mister Maier gets the following data:
```{r}
x_flensburg <- c(2e3, 12e3, 17e3, 12e3, 200e3)
n_flensburg <- length(x_flensburg)
```

#### Calculation of which means is meaningful in this case?

```{r}
# Arithmetic mean
mean(x_flensburg)

# Geometric mean
prod(x_flensburg)^(1 / length(x_flensburg))
```

A telex corrects the value for the last order to 2 000 instead of 200 000.

#### Based on the corrected data calculate arithmetic mean and standard deviation of order values.

```{r}
x_flensburg[length(x_flensburg)] <- 2000

x_bar_flensburg <- mean(x_flensburg)
x_bar_flensburg
# IMPORTANT: Don't use sd as it computes the sample 
# standard deviation instead of the empirical 
# standard deviation
sd_flensburg <- sqrt(1 / n_flensburg * sum((x_flensburg - x_bar_flensburg)^2))
var_flensburg <- sd_flensburg^2
```

#### Compare the deviation of order value between the New York and the Flensburg store using an appropriate measure.

```{r}
# New York (compare Part B)
var_new_york
# Flensburg
var_flensburg

# New York and Flensburg stores share the same standard deviation 
# and variance.
```

#### Calculate the mean order value for the whole business

```{r}
# Berlin (compare Part A)
total_order_value_berlin <- 100 * 79.5e3

# New York (compare Part B)
total_order_value_new_york <- 95 * 90e3

# Flensburg
total_order_value_flensburg <- sum(x_flensburg)

mean_order_value <- (
  total_order_value_berlin + total_order_value_new_york + total_order_value_flensburg
) / (100 + 95 + 5)
mean_order_value
```

#### Calculate the empirical variance of order value for the whole business.

```{r}
# Berlin variance
n_orders <- c(15, 30, 45, 10)
x_bar_berlin_groups <- c(10e3, 35e3, 100e3, 225e3)
var_berlin <- 1 / n_berlin * sum(n_orders * (x_bar_berlin_groups - x_bar_berlin)^2)
var_berlin

# Values from previous tasks
n_cities <- c(n_berlin, n_new_york, n_flensburg)
x_bar_cities <- c(x_bar_berlin, x_bar_new_york, x_bar_flensburg)
var_cities <- c(var_berlin, var_new_york, var_flensburg)

# Total variance
var_all <- 1 / sum(n_cities) * (
  sum(n_cities * var_cities) + sum(n_cities * (x_bar_cities - mean(x_bar_cities))^2)
)
```

#### Which percentage of total variance is generated by considering each store separately?

```{r}
1 / sum(n_cities) * sum(n_cities * (x_bar_cities - mean(x_bar_cities))^2) / var_all
```

## Exercise B2

Every week, the production manager of a company that manufactures tools has to work out a machine allocation plan. For this purpose he needs all orders with the respective quantities [in thousands of pieces]. The computer of the company provides him with the data of this week in the following graphic representation:

```{r, echo=FALSE}
x_dot <- c(15, 25, 30, 35, 45)
y_dot <- c(4, 5, 7, 8, 12)
x_point <- x_dot
y_point <- c(0, 4, 5, 7, 8)
tbl <- tibble(x_dot = x_dot, y_dot = y_dot, x_point = x_point, y_point = y_point)

ggplot(tbl) +
  geom_segment(aes(x = x_point, xend = x_point, y = y_point, yend = y_dot), linetype = "dashed") +
  geom_segment(aes(x = x_point, xend = lead(x_point), y = y_dot, yend = y_dot)) +
  geom_segment(x = max(x_point), xend = 50, y = 12, yend = 12) +
  geom_segment(x = 0, xend = 15, y = 0, yend = 0) +
  geom_point(aes(x = x_dot, y = y_dot), shape = 19, size = 3) +
  geom_point(aes(x = x_point, y = y_point), shape = 1, size = 3) +
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, by = 5)) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, by = 2)) +
  labs(x = "x", y = "H(x)")
```

### What is the statistical feature? How is it scaled?

### Determine absolute and relative frequencies.

### Determine (without calculating) the value of the arithmetic mean and the median. Justify your answer briefly.

### Determine the empirical variance.

### Group order quantities in the following groups: 0 - 20 | 20 - 40 | 40 - 60

### Determine the value of the arithmetic mean and the empirical variance from the grouped data.

### Explain possible differences of the parameter values between original and grouped data

The production manager wants to distribute the production of the ordered tools to the three machines A, B and C.

* Machine A is to produce exactly 40% of this week's order volume.
* For technical reasons, the quantity produced by machine B depends on the production quantity of machine A and is only 25% of the quantity produced by machine A.
* The rest of the order volume is to be produced by machine C.

### How many tools are produced by machine A?

### What percentage of the total order volume is produced by machine B?

### How many tools are produced by machine C?

## Exercise B3

As part of a medical examination, 200 people were asked how many cups of coffee they drink per week. The result of this examination is shown in the following graph:

```{r, echo=FALSE, warning=FALSE}
x <- c(0, 5, 10, 20, 40)
y <- c(0, 0.05, 0.25, 0.7, 1)
tbl <- tibble(x = x, y = y)

ggplot(tbl, aes(x = x, y = y)) +
  geom_point() +
  geom_line() +
  geom_segment(aes(x = 0, xend = x, y = y, yend = y), linetype = "dashed") +
  geom_segment(aes(x = x, xend = x, y = 0, yend = y), linetype = "dashed") +
  scale_x_continuous(limits = c(0, 45), breaks = seq(0, 40, by = 5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1), expand = expansion(c(0, 0.05))) +
  labs(x = "x", y = "F(x)")
```

### What do you call this graphic representation? What was assumed about the distribution of the data?

### Draw the absolute frequency distribution. Which principle should be observed?

### Determine the arithmetic mean and median from the data.

### Interpret both values in terms of how the data is distributed.

### Indicate the percentage of people who drink at least 20 cups of coffee per week.

Let's assume that when you look through the data again, you find that the frequencies of the 3rd group and the 4th group have been swapped.

### Does this change the value of the median?

## Exercise B4

An arithmetic mean of $\bar{x} = 2813$ was determined from $n = 78$ observations. In addition to these observations, an observation of 2100 was added.

### What is the arithmetic mean now?

```{r}
(1 / 79) * (78 * 2813 + 2100)
```

## Exercise B5

Bert is a computer freak. In order to avoid losing touch with the computer world, he likes to be kept up to date with all the latest developments. For this purpose he connected his computer (all day long) to a so-called internet server, for which he has to pay hefty telephone charges. Due to a tariff change of the telephone company, Bert now fears that he can no longer afford this luxury, which is exactly the case when he has to pay more than 6 EUR per hour on average for the telephone connection.

```{r, echo=FALSE}
tbl <- tribble(
  ~tariff,            ~time,           ~seconds_per_cost,
  "Night Tariff",     "01:00 - 06:00", 120,
  "Morning Tariff",   "06:00 - 12:00", 40,
  "Afternoon Tariff", "12:00 - 19:00", 30,
  "Evening Tariff",   "19:00 - 01:00", 60
)
names(tbl) <- c("Tariff", "Time from ... up to less than ...", "Talk time in seconds per cost unit of 8 Cent")

kable(tbl)
```

### Calculate for the new tariff how much the telephone connection costs on average in EUR per hour. So can Bert still afford his hobby in the future?

## Exercise B6

The following empirical distribution function $F(x)$ was obtained for the monthly gross income [in 1000 EUR] of the 2000 employees of a company:

```{r, echo=FALSE}
x <- c(1, 3, 4, 5, 6, 8, 10)
y <- c(0, 0.15, 0.4, 0.7, 0.85, 0.95, 1)
tbl <- tibble(x = x, y = y)

ggplot(tbl, aes(x = x, y = y)) +
  geom_point() +
  geom_line() +
  geom_segment(aes(x = x, xend = x, y = 0, yend = y), linetype = "dashed") +
  geom_segment(aes(x = 0, xend = x, y = y, yend = y), linetype = "dashed") +
  geom_segment(x = 10, xend = 10.5, y = 1, yend = 1) +
  labs(x = "x", y = "F(x)") +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = 0:10) +
  scale_y_continuous(limits = c(0, 1), expand = expansion(c(0, 0.05)), breaks = seq(0, 1, by = 0.1))
```

### What is the name of the statistical feature examined here and how is it scaled?

### Determine the absolute and the relative frequency distribution of this characteristic

### Visualize the absolute frequency distribution.

### According to your visualization, how big is the percentage of employees who earn exactly 6400 EUR per month?

### Calculate how many percent of the employees earn more than 5800 EUR per month.

### What is the (estimated) monthly gross income of all employees of this company?

### What is the average monthly gross income in this company

#### by the arithmetic mean?

#### by the median?

### Calculate the empirical quartile coefficient of skewness.

### Assess the skewness of the present frequency distribution

#### according to the position rule.

#### according to the empirical quartile coefficient of skewness.

## Exercise B7

Multiple Choice Questions

1. One speaks of the _principle of equal area_ when, in the graphic representation of absolute frequencies of grouped data, the flatness of a block is proportional to the absolute frequency to be represented.

  + True.

2. The frequency distributions of two cardinally scaled features do not have to be identical even if the two features have the same arithmetic mean and the same variance.

  + True.

3. Statistical quantities can be classified with respect to their value range.

  + True.

4. The empirical distribution function is not only defined for values that were also observed.

  + False? Empirical means observed. One could argue that she could use aggregated values instead of bare observations.

5. The empirical quartile distance of an ordered data set is a measure of variability which is outlier-sensitive.

  + False.  

6. The step heights of an empirical (relative) distribution function correspond to the relative cumulative frequencies.

  + False, they correspond to the relative frequencies.

7. The coefficient of variation is a dimensionless measure.

  + True.

8. If: $h(a_j) = n$ , then the underlying statistical characteristic has only one possibility of realization. 

  + False. Other possible realizations were just not observed, but may surely exist.

9. A train travels $97 \, \mathrm{km}$ at a speed of $64 \frac{\mathrm{km}}{\mathrm{h}}$ and another $97 \, \mathrm{km}$ at a speed of $192 \frac{\mathrm{km}}{\mathrm{h}}$ . Then the average speed for the total distance is 96km/h.

  + True. $\frac{97 \, \mathrm{km} + 97 \, \mathrm{km}}{\frac{1}{64} \frac{\mathrm{h}}{\mathrm{km}} \cdot 97 \, \mathrm{km} + \frac{1}{192} \frac{\mathrm{h}}{\mathrm{km}} \cdot 97 \, \mathrm{km}} = 96 \frac{\mathrm{km}}{\mathrm{h}}$.

10. A statistical characteristic is called continuous if its range of values is countably infinite.

  + False. Its range must be uncountably infinite.

## Exercise B8

The gross monthly wage of the head of household was determined for 20 households. The following values [in EUR] resulted:

```{r}
x <- c(2400, 3300, 2400, 2300, 2500, 1700, 2100, 3100, 2900, 1800, 3900, 2100, 2900, 1500, 2700, 2200, 2000, 2500, 2600, 2200)
```

### Determine the average monthly gross earnings.

### Calculate the empirical variance and standard deviation of the 20 gross wages.

### Determine the variation coefficient. What is this coefficient for?

## Exercise B9

In Metropolis, 4000 residents with $50 \, \mathrm{m}^2$ living space were asked about their rent levels. In order to get a better overview, the rent amounts were divided into classes.

The following histogram resulted:

```{r, echo=FALSE}
x_left <- c(800, 1000, 1400, 1500)
x_right <- c(1000, 1400, 1500, 2000)
y <- c(0.0015, 0.001, 0.002, 0.0002)
tbl <- tibble(x_left = x_left, x_right = x_right, y = y)

ggplot(tbl) +
  geom_rect(aes(xmin = x_left, xmax = x_right, ymin = 0, ymax = y), fill = NA, color = "black") +
  scale_x_continuous(breaks = seq(800, 2000, by = 200)) +
  scale_y_continuous(breaks = c(0, 0.0002, 0.0006, 0.001, 0.0015, 0.002), minor_breaks = NULL) +
  labs(x = "x", y = "f_j / b_j")
```

### Calculate the values of the empirical distribution function at the locations:

#### x = 1000

#### x = 1350

#### x = 1950

### Draw the empirical distribution function.

### Calculate the arithmetic mean.

### Calculate:

#### 0.25-Quantile

#### Median

#### 0.75 Quantile

#### Draw a box plot.

### Calculate the empirical quartile coefficient of skewness g_0.25.

### Which statement regarding the skewness of the present distribution can be derived from

#### the calculated skew coefficient g_0.25

#### the position rule (comparison of median and arithmetic mean) 

### Give a short explanation for the different results you have received in the previous task.

## Exercise B10

Tony is editor-in-chief of the automotive magazine _Brummy!_. In the anniversary edition he would like to present to his readers a statistical analysis of the cubic capacity [in cm3] of the cars tested so far.
For this purpose he had the following absolute distribution function prepared by his trainee:

```{r, echo=FALSE}
x <- c(1400, 1600, 1800, 2000, 2500, 3000, 5000)
y_point <- c(0, 3, 7, 10, 16, 18, 19)
y_dot <- c(3, 7, 10, 16, 18, 19, 20)
tbl <- tibble(x = x, y_point = y_point, y_dot = y_dot)

ggplot(tbl) +
  geom_point(aes(x = x, y = y_point), shape = 1, size = 3) +
  geom_point(aes(x = x, y = y_dot), shape = 19, size = 3) +
  geom_segment(aes(x = 0, xend = x, y = y_dot, yend = y_dot), linetype = "dashed") +
  geom_segment(aes(x = x, xend = x, y = 0, yend = y_dot), linetype = "dashed") +
  geom_segment(aes(x = x, xend = lead(x), y = y_dot, yend = y_dot)) +
  geom_segment(x = 5000, xend = 5500, y = 20, yend = 20) +
  scale_x_continuous(breaks = seq(0, 5000, by = 1000), limits = c(0, 5000), expand = expansion(c(0, 0.05)), minor_breaks = seq(0, 5000, by = 200)) +
  scale_y_continuous(breaks = seq(0, 20, by = 2), minor_breaks = 0:20, limits = c(0, 20), expand = expansion(c(0, 0.05))) +
  labs(x = "x", y = "H(x)")
```

### Part A

#### What is the statistical characteristic and how is it scaled?

#### Determine the absolute frequencies from the distribution function in tabular form.

#### Calculate the arithmetic mean and the empirical standard deviation.

#### Transform the parameter values calculated in 3) into liters. The following applies: 1 liter = 1dm^3

### Part B

Tony wants to divide the data into three groups:

* Middle Class: $\text{from} \, 1400 \, \mathrm{cm}^3 \, \text{to under} \, 2000 \, \mathrm{cm}^3$
* Middle Class: $\text{from} \, 2000 \, \mathrm{cm}^3 \, \text{to under} \, 3000 \, \mathrm{cm}^3$
* Middle Class: $\text{from} \, 3000 \, \mathrm{cm}^3 \, \text{to under} \, 6000 \, \mathrm{cm}^3$

#### Determine the absolute and relative frequencies of the grouped data in tabular form.

#### Calculate the arithmetic mean of the grouped data.

#### Display the absolute frequency distribution of the grouped data.

#### Determine the absolute cumulative and relative cumulative frequencies of the grouped data in a table and display the relative distribution function graphically. What assumptions are made about the distribution of data within the individual groups?

#### Calculate the lower and upper quartiles and the median from the grouped data.

#### Create a modified box plot for the grouped data.

<!--chapter:end:02_B_Empirical_Parameters.Rmd-->

# Concentration Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Exercise C1

## Exercise C2

## Exercise C3

## Exercise C4

## Exercise C5

## Exercise C6

## Exercise C7

## Exercise C8

## Exercise C9

## Exercise C10

<!--chapter:end:03_C_Concentration_Analysis.Rmd-->

# Multivariate Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Exercise D1

In a survey, 200 father-son pairs were asked about their voting behavior in the last election. Create a contingency table using the following information:

1. The sons chose parties $A$, $B$ and $C$ in the ratio $2:1:2$.
2. Half of the fathers who voted for Party $B$ must face the fact that their sons prefer Party $A$.
3. 5% of the sons interviewed - like their fathers - chose party A.
4. Party B won 40% of all votes.
5. 15% of all fathers gave their vote to Party C.
6. 20 sons whose fathers voted for Party A chose Party B.
7. Only 30 fathers voted for the same party as their sons.

## Exercise D2

In the context of developing a marketing strategy for their latest product, the marketing department of a company conducted a survey in the hope of obtaining information about the most promising packaging color. For this purpose, a total of 600 units of the product were offered on three different test markets. Several employees have so far collected the following information about the pieces sold in the test markets:

1. The number of units sold on test market I was in line with the packaging color (black, red, white) in the ratio 1:3:1.
2. 30% of the pieces sold on the Test Market II were packed in black.
3. 25% of all offered pieces were not sold.
4. As many black-packaged pieces were sold as white-packaged pieces.
5. 175 pieces were sold on Test Market III.
6. The number of pieces sold on Test Market II represented 75% of the total number of red wrapped pieces sold on all Test Markets.
7. On Test Market I, 50 fewer pieces were sold than on Test Market III.
8. 35% of the red wrapped pieces sold were sold on Test Market III.

### Present the data obtained from the examination in tabular form in a contingency table.

### Determine the sales shares of the individual packaging colors in the various test markets.

### Based on the results obtained in 2), which packaging color is the marketing department likely to choose?

## Exercise D3

A survey was carried out in which the interviewed persons were asked to give information about their love for animals (yes/no). The respondents were divided into three categories (children, teenagers, adults).

The survey produced the following results:
1. Of the 400 children questioned, 80% are fond of animals.
2. 200 adults were interviewed.
3. The number of animal-loving adults was 140 less than the number of animal-loving children.
4. Altogether 80% (=800 persons) were animal-loving.

### What is the number of non-animal-loving adults?

### What percentage of respondents are not adults?

### How large is the proportion of animal-loving teenagers among those surveyed?

### How large is the proportion of teenagers among animal lovers?

### Do children love animals more than adults?

## Exercise D4

An investigation by the public health department in X-City revealed the following findings about infections with the virus Z in risk groups A, B and C:

```{r, echo=FALSE}
group <- LETTERS[1:3]
infected <- c(18, 5, 17)
not_infected <- c(42, 31, 37)
tbl <- tibble(group = group, infected = infected, not_infected = not_infected)
names(tbl) <- c("group", "infected", "not infected")

kable(tbl)
```

The realizations of the feature _Group_ are denoted by $a_i, i \in \{A, B, C\}$, whereas the realizations of the feature _Infected_ are denoted by $b_j \, , j \in \{1, 2\}$.

###  Is the frequency f(a_i|b_j) more interesting for any member of one of these risk groups than the frequency f(b_j|a_i)? Explain.

### Is the infection level for group B the lowest in this investigation?

### Is the statement correct that in group B the fewest non-infected people were found?

### Is it correct to say that the proportion of non-infected people is smallest in group C?

## Exercise D5

### Part A

After intensive research work, the Birnwiese Fruit Growers' Association would like to bring two new apple varieties Grüner Bitterer (G) and Kleiner Saurer (K) to the market. To test their market chances, the association will supply the new varieties together with the two already established varieties Edler Gelber (E) and Rotbaeckchen (R) to three canteen kitchens (a school kitchen S, a mensa M and a company canteen C) in a total quantity of all apple varieties of 200 kg.

1. S receives 10 kg from variety (R).
2. From (E) a total of 70 kg is delivered.
3. A total of 30 kg is delivered by (G).
4. The delivered quantities of S and M together make up the delivered quantity of C.
5. The quantity delivered from (R) to M makes up 10% of the total quantity delivered to C.
6. From (E) twice as much is delivered to S as from (E) to M.
7. C receives four times as much from (E) as M receives from (E).
8. In total, twice as much is delivered to M as from (E) to S.
9. From (G) four times as much is delivered to S as is delivered from (G) to M and C respectively.
10. From (G) half as much is delivered to S as from (R) to C.

#### What are the names of the statistical features examined here and how are they scaled?

#### Display the common frequency distribution of these two characteristics in tabular form.

#### Are the two characteristics independent of each other in terms of the data available? Justify your answer by means of a formal calculation. 

#### Which of the three canteen kitchens gets the smallest share of new apple varieties in relation to the total quantity of apples delivered to them? Justify your answer with the help of a formal calculation.

### Part B

According to the fruit growers' association, hardly any maggots should be present in the two new apple varieties. The canteen chef distrusts these promises and examines half of the apples of the new varieties delivered to him for maggots ($6$ apples $\approx$ 1 kg). He receives the following result:

```{r}
maggots <- 0:3
rel_frequency <- c(0.2, 0.4, 0.3, 0.1)
tbl <- tibble(maggots = maggots, rel_frequency = rel_frequency)
names(tbl) <- c("Number of maggots per apple", "Relative frequency")

kable(tbl)
```

#### What is the name of the statistical feature examined here and how is it scaled?

#### Determine the absolute frequencies as well as the absolute and relative cumulative frequencies of the statistical feature in a table.

#### Display the relative distribution function.

#### Determine **graphically** the values of a so-called _5 number summary_ and draw the corresponding box-plot.

## Exercise D6

### Part A

The standard of cleanliness in youth hostels often leaves a lot to be desired. A group of statistics students would therefore like to conduct a study in four different regions (A, B, C and D) of the country to get an idea of the standard of cleanliness (good, moderate, poor) of some of the youth hostels located there. A total of 40 youth hostels were inspected with the following results:

1. In region D there are 2 hostels of moderate standard.
2. A total of 14 hostels in Region C and 6 hostels in Region A were checked.
3. In 12 hostels checked, the standard is moderate.
4. In Region C, 8 hostels have a good standard.
5. In Region C, there are twice as many hostels with a moderate standard as those with a poor standard.
6. The hostels in Region C with a poor standard make up 10% of the hostels with a good standard of all regions.
7. In Region A, there are four times as many hostels of moderate standard as there are hostels of good or bad standard in each region.
8. The number of moderate-standard hostels in Region A is half the number of good-standard hostels in Region D.
9. There are four times as many hostels of moderate standard in all regions together as there are hostels of poor standard in Region B.

#### What are the names of the statistical features examined here and how are they scaled?

#### Display the common frequency distribution of these two characteristics in tabular form.

#### Are the two characteristics independent of each other in terms of the data available? Justify your answer by means of a formal calculation.

#### In which region is the cleanliness (of hostels with at least moderate standard) greatest? Justify your answer by means of a formal calculation.

### Part B

After the results of this study were announced, all the managers of those hostels that were certified as having poor standards of cleanliness unanimously assured that a thorough remedy of the deficiencies complained about would be provided as soon as possible. After some time, 20 rooms in each of the affected hostels were checked again for their standard of cleanliness:

```{r, echo=FALSE}
x <- c("[0; 2)", "[2; 5)", "[5; 10)", "[10; 12]")
rel_frequency <- c(0.2, 0.45, 0.25, 0.1)
tbl <- tibble(x = x, rel_frequency = rel_frequency)
names(tbl) <- c("Nuber of Cleanliness Deficiencies per Room", "Relative Frequency")
```

#### What is the name of the statistical feature under investigation here and how is it scaled?

#### Determine the absolute frequencies as well as the absolute and relative cumulated frequencies of the statistical feature in a table.

#### Display the (relative) distribution function graphically.

#### Calculate and graphically determine the values of a so-called 5-number summary and draw the corresponding box plot.

## Exercise D7

### Part A

### Part B

<!--chapter:end:04_D_Multivariate_Analysis.Rmd-->

# Correlation Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Exercise E1

A group of science students wrote an exam in math and an exam in physics. From 10 randomly selected students the results (in points) are summarized in the following table:

```{r}
student <- 1:10
maths <- c(14, 8, 10, 20, 12, 8, 12, 20, 16, 20)
physics <- c(10, 12, 12, 8, 12, 10, 10, 12, 16, 8)
tbl <- tibble(student = student, maths = maths, physics = physics)
names(tbl) <- c("Student", "Maths", "Physics")

kable(tbl)
```

### Part A

#### Calculate the correlation coefficient according to Bravais-Pearson.

### Part B

#### Calculate the (corrected) rank correlation coefficient according to Kendall.

### Part C

Each exam is considered passed if the candidate has achieved at least 12 points.

#### To measure the correlation between the exam result in physics (passed, failed) and the exam result in mathematics (passed, failed), display the exam result of the 10 students in a contingency table.

#### Determine the $\chi^2$ value.

#### Determine the contingency coefficient $K$.

### Part D

#### Give an explanation for the different values of $r$, $\phi$ and $K$.

## Exercise E2

15 apprentices, who participate in a series of sex education lessons as part of their vocational school education, were asked to mark on a five-point scale (1: unimportant, 5: very important) the importance they felt sex education in general should have. In addition, the interviewed apprentices were asked to mark on a three-point scale (1: not at all, 3: a lot) the extent to which the sex education book used in class could contribute to a reasonable sex education. The result is recorded in the following frequency table (rating of important from left to right, rating of book from top to bottom):

```{r, echo=FALSE}
m <- matrix(c("-", 4, 1, "-", "-", "-", "-", 5, 1, "-", 2, "-", 1, "-", 1), nrow = 3, byrow = TRUE)
kable(m)
```

### To indicate the extent to which the two ratings are correlated, calculate the rank correlation coefficient according to Kendall.

## Exercise E3

### Part A

The characteristics of written and oral grades in economics were assessed in 9 exam candidates, with only whole grades from 1.0 to 5.0 being awarded. The result of the survey is shown in the following matrix:

```{r, echo=FALSE}
m <- matrix(c(1, "-", "-", "-", "-", 1, 1, 1, "-", "-", 1, "-", "-", 1, 1, "-", "-", "-", 1, "-", "-", "-", "-", 1, "-"), ncol = 5, byrow = TRUE)
kable(m)
```

#### Calculate a suitable coefficient as a measure of the strength of the relationship between the two characteristics.

### Part B

The characteristics preparation time [in weeks] and number of exams taken were recorded for 10 exam candidates. The result of the survey is shown in the following matrix:

```{r, echo=FALSE}
m <- matrix(c("-", "-", "-", 1, 1, 2, 1, "-", "-", "-", "-", "-", "-", "-", "-", 2, 1, "-", "-", "-", "-", "-", "-", 1, 1), nrow = 5, byrow = TRUE)
kable(m)
```

#### Calculate a suitable coefficient as a measure of the strength of the relationship between the two characteristics.

## Exercise E4

At the University A, those responsible for Statistics I and II exams are interested in using a coefficient to estimate how strong the formal relationship between the two exam results (measured in points) is. For this purpose, 5 students are randomly selected from the total number of students who have written these exams. In the following table their compiled point values are listed:

```{r, echo=FALSE}
student <- 1:5
stat_1 <- c(20, 30, 50, 70, 80)
stat_2 <- c(70, 40, 10, 40, 70)
tbl <- tibble(student = student, stat_1 = stat_1, stat_2 = stat_2)
names(tbl) <- c("Student", "Statistics 1", "Statistics 2")

kable(tbl)
```

### Part A

#### Draw the data into a scatterplot.

#### Calculate the appropriate coefficient for this problem.

#### Interpret the result.

#### What does this coefficient measure and what does it not measure.

### Part B

Assign ranks to the (cardinal) data.

#### Calculate the adequate coefficient.

#### Interpret the result.

## Exercise E5

A psychologist investigates different work groups in a company to see if there is a connection between the attractiveness of a work group (cohesion and well-being of its members) and the variance of its work performance.

Group attractiveness was measured using an ordinal scale from 1 = very high to 6 = very low. The variance in performance refers to the number of products completed by a work group on different days. The study produced the following result: 

```{r, echo=FALSE}
group <- LETTERS[1:7]
attractiveness <- c(1:2, 2:6)
variance <- c(32, 20, 40, 32, 64, 68, 60)
tbl <- tibble(group = group, attractiveness = attractiveness, variance = variance)
names(tbl) <- c("Work Group", "Attractiveness", "Variance in Performance")

kable(tbl)
```

### What is the statistical feature in this study?

### Calculate a suitable correlation coefficient.

### Interpret the calculated sign of this coefficient.

### Does the calculated value of this coefficient provide information about a cause-and-effect relationship between the two characteristics examined?

<!--chapter:end:05_E_Correlation_Analysis.Rmd-->

# Combinatorics and Algebra of Events

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

## Exercise F1

Three doctors work in the emergency room of a hospital: N, O and P. Since the division of weekend services (Saturday and Sunday) is very difficult, the three doctors decide to conduct a random experiment to determine the division. Three slips of paper with the first letters of their names (N, O and P) are placed in an urn. Two pieces of paper are then drawn at random from the urn to make the division.

Specify the possible outputs (2-tuples) of this random experiment, if

### with the allocation you want to determine on which day a doctor is on duty (first element of the 2-tuple stands for Saturday), and it should be possible for a doctor to be on duty (i.e. double duty) on both days.

### double duty is possible, but it should not be determined on which day a doctor is on duty.

### double duty is not possible, but it should be determined on which day a doctor is on duty.

### double duty is not possible and it should not be determined on which day a doctor is on duty.

## Exercise F2

### Part A

At the beginning of the semester, a group of students would like to offer a 5-day orientation tour of the university for first-year students. There are 5 volunteer members available.

#### How many different ways are there to give each member a weekday (Monday to Friday)?

Since Klaus has fallen ill, Karl will lead a tour twice.

#### How many different options are there for scheduling the tours by name?

Karl suddenly insists that he does not want to lead a tour on two consecutive days.

#### How many different options are there for scheduling the tours by name now?

### Part B

In two urns there are 6 different colored balls in each of the two urns, whereby the color composition of the balls in both urns is the same. Two balls are drawn at random from each urn without putting them back.

How many possibilities are there,

#### to draw a total of 2 balls from both urns?

#### to get two pairs of balls of the same color?

#### to get four different colored balls?

#### to get exactly one pair of balls of the same color?

## Multiple Choice

Are the following statements right or wrong?

### Be $A$, $B$, $C$ $\subset \omega$ events. The event: "At least two of these events occur" can be represented graphically as follows:

### From $n$ objects $r$ elements shall be selected. Then the number of choices for a variation with repetition can not be smaller than for a variation without repetition.

### The certain event has the probability one. But an event with a probability of one is only "almost certain".

### Two events, which have no common elements, are disjoint.

### Using the objective concept of probability, the individual case is considered, but not the series ("long run").

### For two events $A$ and $B$ applies: $P(A \cap B) \le P(A) \le P(A \cup B) \le P(A) + P(B)$.

### If three events $A$, $B$ and $C$ are stochastically independent in pairs, then there is valid: $P(A \cap B \cap C) = P(A) \cdot P(B) \cdot P(C)$.

### If three events $A$, $B$ and $C$ are completely disjoint [i.e. $(A \cap B \cap C) = \emptyset$], then they are also disjoint in pairs [i.e. $(A \cap B) = \emptyset$, $(B \cap C) = \emptyset$ and $(C \cap A) = \emptyset$].

### Let $A$ and $B$ be events (with: $0 < P(A) < 1$ and $0 < P(B) < 1$). If $P(A \cap B) = 0$, then $A$ and $B$ are stochastically independent.

### The probability of throwing a non-ideal die twice with a maximum of a "6" is equal to the probability of not throwing a "6" at least once.

## Exercise F4

### Part A

Three different special prizes are raffled off in a game show. Three urns are available for the raffle, each containing six tickets with the names of the six players. One ticket is drawn from each urn.

#### How many possibilities are there to draw a ticket from each of the three urns?

#### How many possibilities are there to draw a ticket from each of the three urns, whereby no name is drawn several times?

### Part B

After the game show, the six candidates (who are all looking for comfort) are taken to their hotel in two cars of different comfort. One car can accommodate up to four candidates. 

#### How many possibilities are there to accommodate the six candidates in the two cars?

### Part C

The publishing house "Schöne Kinderwelt" would like to publish a fairy tale book. The publishing house owns the rights to 6 fairy tales by the Brothers Grimm and 4 fairy tales by Hans Christian Andersen as well as the rights to 2 fairy tales by other authors. For reasons of cost, only 7 fairy tales should be selected.

#### How many possibilities are there to select the 7 fairy tales?

#### How many possibilities remain if 3 fairy tales by the Grimm brothers, 3 by Hans Christian Andersen and 1 fairy tale by other authors are to be selected?

#### How many possibilities are there to arrange the 7 selected fairy tales in a book?

## Exercise F5

In a lottery there are 10 winners. The following cars are available as main prizes:
5 cars from "Idua" | 3 cars from "WMB" | 2 cars from "WV".

### How many ways are there to distribute the cars to the winners?

In another lottery, six companies each provide one car, but the cars differ considerably in price and quality. Only three winners have declared their claim to the prize.

### How many possibilities are there to distribute the cars among the three winners, if everyone should only get one car at a time?

## Exercise F6

Of three events $A$, $B$, $C$ occur:

1. only $A$.
2. exactly one.
3. at most one.
4. at least one.
5. exactly two.
6. at least two.
7. at least one not.
8. at least two not.

### Use the event operations to formally represent the searched events and draw the corresponding Venn diagram for each.

## Exercise F7

In a small town there are only the two daily newspapers $Z1$ and $Z2$. 60% of the inhabitants read $Z1$ and 80% of the inhabitants read $Z2$. Neither newspaper is read by 10% of the residents. A person is chosen at random. What is the probability that this person

1. reads both newspapers?
2. reads $Z1$, but not $Z2$?
3. reads $Z2$ if she is not a Z1 reader?
4. reads at most one newspaper?
5. does not read $Z1$?

## Exercise F8

In an apartment in an old building with exterior walls and an outdated electrical system, it can happen that the electricity fails or the water supply freezes. Both circumstances occur independently, but depend on the season. Thus, of course, the water only freezes when it is winter, with a probability of 80%. However, even if it is not winter, there is a 40% probability that the electricity will fail, which is the same probability that the electricity will not fail when it is winter.
Assume that the winter season accounts for 30% of the total season.

### Use event symbols to formalize the probability statements mentioned in the text.

### With what probability must one expect the water supply to freeze?

### With what probability must one expect a power failure?

### With what probability must one expect the water supply to freeze and a power failure?

### If the power has already been cut off, with what probability must one expect the water supply to freeze?

### With what probability must one additionally expect a power failure if the water supply is already frozen?

### With what probability must one expect at least one of both problems?

### With what probability must one expect at most one of the two problems?

## Exercise F9

At a small border crossing, two customs officers share the duty exactly halfway. When the first customs officer is on duty, he checks passports with a probability of 20% and searches cars with a probability of 10% independently of a passport check. When the second customs officer is on duty, he checks passports with a probability of 70% and searches cars with a probability of 40% independently of a passport control.

### Define meaningful events based on the text and formalize the specified probabilities.

You drive to this border crossing with a car at any time.

### What is the probability that your car will be searched?

### What is the probability of a passport check?

### What is the probability that you will be searched and your car will be checked?

### If your passport has already been checked, how likely are you to have your car searched?

### If your car has already been searched, how likely are you to face an additional passport check?

### What is the probability that you will have to expect exactly one of both measures?

## Exercise F10

A man is looking for his umbrella, which he has left in a building with a 60% probability. And if it is true, with equal probability in one of the 6 floors.

Use the following events for your solution:

* $B$: "Umbrella is in the building."
* $F$: "Umbrella is in the first 5 floors.
* $S$: "The umbrella is on the 6th floor."

### What is the probability that the umbrella is in the building and on the 6th floor of the building?

### What is the probability that the umbrella is on the 6th floor if you don't know if it is in the building?

### What is the probability that the umbrella is not on the first 5 floors if you know that it is not on the 6th floor?

### What is the probability that the umbrella is not on the first 5 floors?

### The umbrella is not in the first 5 floors. What is the probability that the umbrella is in the building?

## Exercise F11

The Road Traffic Office is conducting a study on the aggression behavior of drivers towards cyclists and comes to the following conclusion:

* A male motorist reacts to a hindrance by a cyclist with a 30% probability by "honking" and (independently of this) with a 20% probability by " verbal abuse". 
* A female motorist reacts to a hindrance caused by a cyclist with a 15% probability by "honking" and (independently of this) with a 10% probability by "verbal abuse". 
* Meeting a man behind the wheel or a woman behind the wheel can be considered equally likely.

### Define corresponding events and assign the probability values mentioned in the text to them.

Suppose you are a cyclist who is hindering a person behind the wheel of a car.

### Will this person honk?

### Will this person insult you?

### Will this person honk the horn and insult you?

### Will this person insult you even if he or she honks already?

## Exercise F12

### Part A

In a horse race exactly seven horses start.

#### How many possibilities are there to divide the seven horses among the ten starting boxes?

#### How many possible finishes are there in places one to seven?

#### How many possibilities are there for the distribution of ranks one to three?

### Part B

Today is a big race day in Hoppewald with three scheduled races. In the individual races the following horses will be at the start:

* Race I : $A$, $B$ and $C$
* Race II : $D$, $E$, $F$, $G$ and $H$
* Race III: $A$, $B$, $C$, $D$, $E$, $F$, $G$ and $H$

The following victory probabilities apply to the individual horses, with the outcomes of the three races being independent of each other:

```{r}
race <- c(rep("I", times = 2), rep("II", times = 4), rep("III", times = 7))
horse <- c(LETTERS[1:2], LETTERS[4:7], LETTERS[1:7])
victory <- c(0.2, 0.4, 0.1, 0.2, 0.3, 0.4, 0.05, 0.1, 0.05, 0.05, 0.1, 0.1, 0.05)
tbl <- tibble(race = race, horse = horse, victory = victory)
names(tbl) <- c("Race", "Horse", "Probability of Victory")

kable(tbl)
```

What are the probabilities of the following events:

#### Horse $C$ wins exactly one of his races.

#### Horse $C$ wins both of his races.

#### Horse $C$ wins at least one of his races.

#### Horse $C$ wins at most one of his races.

#### Horse $C$ wins none of his races.

The horses C and H belong to the same racing stable.

#### What is the probability that this racing team will be able to celebrate a winner today, i.e. to be the winner in at least one race?

## Exercise F13

### Part A

In Skat, an ideal deck of cards (32 cards) is distributed among three players and the "Skat". Each player receives 10 cards, the remaining 2 cards go into the "Skat". The four existing jacks have a particularly important meaning in the Skat game. 

#### What is the probability that two jacks are in the "Skat"?

#### What is the probability that exactly one jack is in the "Skat"?

#### What is the probability that the jack of clubs is in the "Skat"?

### Part B

Two ideal dice are thrown. One cube is red, the other is white. The following events are defined:

* $A$: "The number of the white cube is even."
* $B$: "The number of the red cube is odd."
* $C$: "The sum of the eyes of both cubes is straight."
* $D$: "The sum of the eyes of both cubes is odd."

#### Determine the probabilities of the events $A$, $B$, $C$ and $D$.

#### What elements of $\Omega$ make up the following events?

1. $A \cap B$.
2. $A \cap C$.
3. $B \cap C$.

#### Use _Laplace_'s concept of probability to determine the probabilities of the events

#### Are the events $A$, $B$ and $C$ pairwise stochastically independent? Justify your respective answer with the help of a formal arithmetic operation.

#### Determine the probability of the event: $(A \cap B \cap C)$

#### Are the events $A$, $B$ and $C$ completely stochastically independent? [i.e. $P(A \cap B \cap C) = P(A) \cdot P(B) \cdot P(C)$]

One of the following statements is correct. Indicate which statement is correct:

1. "Completely stochastically independent events are always also stochastically independent in pairs."
2. "Pairwise stochastically independent events are always also completely stochastically independent."

### Part C

In a tennis match, player A wins a single set against player B with probability $p$. The winner of the whole match is the first player to win two sets.

Be $G_i$ : "Player A wins the ith set" with $i \in \{ 1, 2, 3 \}$.

### Depending on p, specify the probability that player A will win the match.

## Exercise F14

A family was at the picnic today: Father Martin, mother Silke, the children Anja and Dirk as well as grandpa Arnold. On the way home the children suddenly notice that grandpa is no longer there. There are three possibilities:

* $A$: Grandpa is already at home and sitting comfortably in his armchair.
* $B$: Grandpa is still at the picnic area and taking his nap.
* $C$: Grandpa has gone to the nearby forest and is looking for mushrooms.

Because of the habits of grandpa, one knows the probabilities for the occurrence of events $A$, $B$ and $C$:

$$P(A) = 0.15, \, P(B) = 0.8, \, P(C) = 0.05 \, .$$

Anja is sent back to the picnic place and Dirk is sent to the edge of the forest to look for the grandfather. If grandpa is at the picnic place, Anja finds him with a probability of 90%. But if he is running around in the forest, Dirk will find him with a probability of only 50%.

### What is the probability that Anja will find the grandpa?

### What is the probability that Dirk will find the grandpa?

### What is the probability that one of the children will find the grandpa?

### What is the probability that the grandpa will be sitting in his armchair when he returns home, in case the children do not find him?

## Exercise F15

A florist has an assortment of tulip bulbs. Of these bulbs 20% make blue tulips, the rest make red tulips. 60% of the bulbs make tulips with smooth leaves, the rest make tulips with pointed leaves. If a tulip is blue, it has smooth leaves in 10% of the cases.

It applies:

* R: "The tulip turns red."
* G: "The tulip will have smooth leaves."

### Calculate the probability that a tulip will have red color **and** smooth leaves.

### Calculate the probability that a tulip will turn red if you know that it will have smooth leaves.

### Calculate the probability that a tulip will have smooth leaves if you know that it will turn red.

### Calculate the probability that a tulip will turn red if you know that it will turn pointed.

### Calculate the probability that a tulip will turn red **or** smooth if you know it will turn red **and** smooth.

### Calculate the probability that a tulip will become blue **or** smooth if you know it will become red **and** pointed.

### Calculate the probability that a tulip will turn red **or** smooth if you know it will turn red **and** pointed.

### Calculate the probability that a tulip will turn red **and** smooth if you know it will turn red **or** smooth.

<!--chapter:end:06_F_Combinatorics_and_Algebra_of_Events.Rmd-->

# Discrete Random Variables and Distributions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

## Prerequisites {-}

```{r}
library(ggplot2)
library(dplyr)
library(knitr)

# Set ggplot2 theme
theme_set(theme_bw())
```

<!--chapter:end:07_G_Discrete_Random_Variables_and_Distributions.Rmd-->

# Continuous Random Variables and Distributions

## Exercise H1

## Exercise H2

## Exercise H3

## Exercise H4

## Exercise H5

## Exercise H6

## Exercise H7

## Exercise H8

## Exercise H9

## Exercise H10

## Exercise H11

## Exercise H12

## Exercise H13

## Exercise H14

## Exercise H15

## Exercise H16

## Exercise H17

<!--chapter:end:08_H_Continuous_Random_Variables_and_Distributions.Rmd-->

